clusters

### 2015-07-10 cluster.rmd
finished clustering - 11 clusters

# library('knitr')
# library('ggplot2')
# library("dplyr")
# library("plyr")
# library('magrittr')
# library('reshape2')
# library("rpart")
# library('car')
# library('e1071')
# library('corrgram')
# library('party')
# library('dendextend')
# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
# opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")

all7c<- read.csv('all7c.csv')[,-1]
all7c$Post.Code<- as.factor(all7c$Post.Code)

@@

```{r final_var, echo=FALSE}

@@

```{r, echo=FALSE}
all7c[,!sapply(all7c, is.factor)] %>% colnames
all7d[,!sapply(all7d, is.factor)] %>% colnames

@@

```{r var_check, echo=FALSE}

@@

```{r clust final, echo=FALSE}
clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
                         cost.mlsto, return.pdol)

@@

#now to standardise
clust<- scale(clust[,-1]) %>% as.data.frame
```

@@

```{r hierarch, echo=FALSE}

@@

x<- dist(clust, method = 'euclidean')
fit<- hclust(x, method = 'ward.D2')
plot(fit)

@@

```{r kmeans, echo=FALSE}

k1<- kmeans(clust %>% select(-groups), 5)
k1.centres<- k1$centers %>% as.data.frame
k1.centres$group<- rownames(k1.centres)
k1.melt<- melt(k1.centres)
k1.melt$group<- k1.melt$group %>% as.factor
ggplot(k1.melt, aes(x=variable, y= value, fill=variable)) + geom_bar(stat='identity', position='dodge')+
        facet_wrap(~group) +
        theme(axis.text.x=element_text(angle=45,hjust=1))

all7c$groups<- k1$cluster
kmeans.plot<- all7c %>% select(inv.mlsto, timespan, no.users, pc.pro, client.totinv,
                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
                         cost.mlsto, return.pdol, groups)
all7c.plot<- ddply(kmeans.plot, .(groups), numcolwise(mean))
all7c.plot$groups<- rownames(all7c.plot)
k2.melt<- melt(all7c.plot)

#summary means of all numeric variables in 5 clusters
clust.summary<- all7c.plot %>% select(-groups) %>% round(2)
tab<- table(all7c$groups) %>% prop.table %>% as.data.frame
bal<- sum(all7c$balance.mlsto)
hrs<-  sum(all7c$hrs.mlsto)
clust.summary$pc.jobs<- round( tab[,2]*100, 1)
clust.summary$bal.pc<- ddply(all7c, .(groups), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
        select(2)
clust.summary$hrs.pc<- ddply(all7c, .(groups), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
        select(2)

clust.summary<- clust.summary[,c(1,12,10,11, 13, 9,2,7,6,3,4,8,5,14:16)]

all7c.med<- all7c.med[, c(2,13,11,12,14,10,3,8,7,4,5,9,6)]

@@

```{r optimise clustnum,echo=FALSE}
library('fpc', lib='C:/Progra~1/R/R-3.2.0/library')
library('vegan', lib = 'C:/Progra~1/R/R-3.2.0/library')

@@

## try out all iffy variables together, eliminate duds
clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
                                      client.totinv, Num.days, hrs.mlsto, pc.pro, mean.peeps,
                                      Num.disc), 2, 20, iter= 100)
plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
# all together, no good
# smaller subset of variables I am most interested in that still works for clustering:

clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
                                      hrs.mlsto
                                      ), 2, 20, iter= 100)
plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)

# 17 clusters
# Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
# then calinksi (index of cluster homogeneity and distance where higher number is better)
# calinski: between cluster distance/within cluster distance
ind.plot<- melt(clust.km$results)
ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
        geom_line( ) +
        theme(axis.text.x=element_text(angle=45,hjust=1))
ggplot(ind.plot %>% filter(Var1 == 'calinski'), aes(x=Var2, y=value, group=1)) + 
        geom_line( )+
        theme(axis.text.x=element_text(angle=45,hjust=1))
# 17 clusters should be good

@@

```{r new var, echo=FALSE}
#new variable for all7c - success
# success= return.pdolxinv.mlsto

all7c<- transform(all7c, suc= inv.mlsto*return.pdol)
all7c$suc %>% summary

all7c %>% arrange(-suc) %>% head

@@

qqPlot(sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)}))
qqPlot((all7c$suc) %>% math.cbrt)

#can't log because of negatives

@@

# 10 clusters
# Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
# then calinksi (index of cluster homogeneity and distance where higher number is better)
# calinski: between cluster distance/within cluster distance
ind.plot<- melt(clust2.km$results)
ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
        geom_line( ) +
        theme(axis.text.x=element_text(angle=45,hjust=1))
ggplot(ind.plot %>% filter(Var1 == 'calinski'), aes(x=Var2, y=value, group=1)) + 
        geom_line( )+
        theme(axis.text.x=element_text(angle=45,hjust=1))

# 10 clusters should be good

# have a look at dendrogram with 10 clusters

@@

rect.hclust(fit, k=10, border='red')

@@

now want 10 clusters, based on k means optimisation, have a look at average characteristics

```{r kmeans, echo=FALSE}

k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 10)

@@

clust.summary<- clust.summary %>% select(inv.mlsto, suc, return.pdol, cost.mlsto, balance.mlsto, hrs.mlsto, hours.perday, timespan, Num.days, Num.disc, no.users, pc.pro, mean.peeps, client.totinv, pc.jobs, bal.pc, hrs.pc)

@@

>>>>

# library('knitr')
# library('ggplot2')
# library("dplyr")
# library("plyr")
# library('magrittr')
# library('reshape2')
# library("rpart")
# library('car')
# library('e1071')
# library('corrgram')
# library('party')
# library('dendextend')
# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
# opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
all7c<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7c.csv')[,-1]
all7c$Post.Code<- as.factor(all7c$Post.Code)

@@

```{r final_var, include=FALSE}

@@

```{r, include=FALSE}
all7c[,!sapply(all7c, is.factor)] %>% colnames
all7d[,!sapply(all7d, is.factor)] %>% colnames

@@

```{r var_check, include=FALSE}

@@

```{r clust final, include=FALSE}
clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
                         cost.mlsto, return.pdol)

@@

#now to standardise
clust<- cbind(scale(clust[,-1]) %>% as.data.frame, 'mlsto'=clust$mlsto)
```

@@

```{r hierarch, include=FALSE}

@@

x<- dist(clust %>% select(-mlsto), method = 'euclidean')
fit<- hclust(x, method = 'ward.D2')
plot(fit)

@@

```{r kmeans, include=FALSE}

# k1<- kmeans(clust %>% select(-groups, -mlsto), 5)
# k1.centres<- k1$centers %>% as.data.frame
# k1.centres$group<- rownames(k1.centres)
# k1.melt<- melt(k1.centres)
# k1.melt$group<- k1.melt$group %>% as.factor
# ggplot(k1.melt, aes(x=variable, y= value, fill=variable)) + geom_bar(stat='identity', position='dodge')+
#         facet_wrap(~group) +
#         theme(axis.text.x=element_text(angle=45,hjust=1))
# 
# #try plotting means of actual values
# 
# all7c$groups<- k1$cluster
# kmeans.plot<- all7c %>% select(inv.mlsto, timespan, no.users, pc.pro, client.totinv,
#                          Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
#                          cost.mlsto, return.pdol, groups)
# all7c.plot<- ddply(kmeans.plot, .(groups), numcolwise(mean))
# all7c.plot$groups<- rownames(all7c.plot)
# k2.melt<- melt(all7c.plot)
# 
# #summary means of all numeric variables in 5 clusters
# clust.summary<- all7c.plot %>% select(-groups) %>% round(2)
# tab<- table(all7c$groups) %>% prop.table %>% as.data.frame
# bal<- sum(all7c$balance.mlsto)
# hrs<-  sum(all7c$hrs.mlsto)
# clust.summary$pc.jobs<- round( tab[,2]*100, 1)
# clust.summary$bal.pc<- ddply(all7c, .(groups), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
#         select(2)
# clust.summary$hrs.pc<- ddply(all7c, .(groups), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
#         select(2)
# 
# clust.summary<- clust.summary[,c(1,12,10,11, 13, 9,2,7,6,3,4,8,5,14:16)]
# 
# # add medians
# all7c.med<- ddply(kmeans.plot, .(groups), numcolwise(median))
# 
# all7c.med<- all7c.med[, c(2,13,11,12,14,10,3,8,7,4,5,9,6)]

@@

```{r optimise clustnum,include=FALSE}
library('fpc', lib='C:/Progra~1/R/R-3.2.0/library')
library('vegan', lib = 'C:/Progra~1/R/R-3.2.0/library')

@@

## try out all iffy variables together, eliminate duds
# clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
#                                       client.totinv, Num.days, hrs.mlsto, pc.pro, mean.peeps,
#                                       Num.disc), 2, 20, iter= 100)
# plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
# # all together, no good
# # smaller subset of variables I am most interested in that still works for clustering:
# 
# clust.km<- cascadeKM(clust %>% select(inv.mlsto, balance.mlsto, return.pdol,
#                                       hrs.mlsto
#                                       ), 2, 20, iter= 100)
# plot(clust.km, sortg = TRUE, grpmts.plot = TRUE)
# 
# # 17 clusters
# # Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
# # then calinksi (index of cluster homogeneity and distance where higher number is better)
# # calinski: between cluster distance/within cluster distance
# ind.plot<- melt(clust.km$results)
# ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
#         geom_line( ) +
#         theme(axis.text.x=element_text(angle=45,hjust=1))
# ggplot(ind.plot %>% filter(Var1 == 'calinski'), aes(x=Var2, y=value, group=1)) + 
#         geom_line( )+
#         theme(axis.text.x=element_text(angle=45,hjust=1))
# 17 clusters should be good

@@

```{r new var, include=FALSE}
#new variable for all7c - success
# success= return.pdolxinv.mlsto

all7c<- transform(all7c, suc= inv.mlsto*return.pdol)
# all7c$suc %>% summary

# all7c %>% arrange(-suc) %>% head

@@

qqPlot(sapply(all7c$suc, function(x){sign(x)*log(abs(x) + 1)}))

@@

# 11 clusters
# Cluster cohesion: create plots of SSE index (distance within clusters - sum of squared error) 
# then calinksi (index of cluster homogeneity and distance where higher number is better)
# calinski: between cluster distance/within cluster distance
ind.plot<- melt(clust.km$results)
ggplot(ind.plot %>% filter(Var1 == 'SSE'), aes(x=Var2, y=value, group=1)) + 
        geom_line( ) +
        theme(axis.text.x=element_text(angle=45,hjust=1))
ggplot(ind.plot %>% filter(Var1 == 'calinski'), aes(x=Var2, y=value, group=1)) + 
        geom_line( )+
        theme(axis.text.x=element_text(angle=45,hjust=1))

# 11 clusters should be good

# have a look at dendrogram with 11 clusters

@@

rect.hclust(fit, k=11, border='red')

@@

now want 11 clusters, based on k means optimisation, have a look at average characteristics

```{r, echo=FALSE}

k1<- kmeans(clust %>% select(inv.mlsto, balance.mlsto, return.pdol, suc, hrs.mlsto), 11)

@@

clust.summary<- clust.summary %>% select(inv.mlsto, suc, return.pdol, cost.mlsto, balance.mlsto,
                                         hrs.mlsto, hours.perday, timespan, Num.days, Num.disc,
                                         no.users, pc.pro, mean.peeps, client.totinv, pc.jobs,
                                         bal.pc, hrs.pc, job.count)

@@

clust.summary<- cbind(all7c.med %>% select(inv.mlsto, suc, return.pdol, cost.mlsto, balance.mlsto,
                                     hrs.mlsto, hours.perday, timespan, Num.days, Num.disc,
                                     no.users, pc.pro, mean.peeps, client.totinv),
                      clust.summary %>% select(pc.jobs,bal.pc, hrs.pc, job.count))

clust.summary$bal<- ddply(all7c, .(groups), summarise, bal= round(sum(balance.mlsto), 1)) %>% 
        select(2)
